### 2011-12-30 joint action and knowing others minds/interacting mindreaders.2.tex
part way through proofing

We are not suggesting that no observer could ever identify the goal of any action she fails to recognise as a means to achieving that goal, of course.
Of course opaque means are not in every case an insurmountable obstacle to goal ascription.
This point is neither novel nor surprising:
opaque means sometimes  deprive interpreters of evidence and so prevent goal ascription.

Some of the most plausibly unique aspects of human cognition depend on our abilities to recognise the goals of novel behaviours involving tools, and of communicative gestures.
If goal ascription is based on entirely observation 
(so that the possibility of interaction is ignored),
the problem of opaque means is likely to arise in both cases.
We have just seen an illustration of how the problem of opaque means arises where tools are used to unfamiliar ends.
Relatedly, it is also likely to arise where actions involve multiple steps that do not form a familiar sequence, can occur in various orders and can be interspersed among other activities;
as in preparing spirit from grain, for example.
%\footnote{
%*cut because 'plausibly unique aspects of human cognition'*
%one example is food preparation in some groups of mountain gorillas
%\citep[p.\ 531]{Byrne:2003wx}.
%*This is slightly tricky because Byrne wants a pure behaviour reading account with no understanding of intention.  
%Stress that this is consistent with supposing that goals are ascribed.
%}

@@

This, then, is the problem of opaque means:
failures to identify to which ends actions are means can impair goal ascription.
The problem is potentially a problem for interpreters given the standard, purely observational models of interpretation.
It is not our intention to suggest that the problem of opaque means is a problem \emph{for models of interpretation}.
Rather, it is a potential problem \emph{for interpreters}.

That the problem of opaque means exists
is the first premise in our argument
that models of interpretation based on pure observation are less powerful than models taking into account the possibility of interaction.
They are less powerful in this sense: 
some routes to knowledge of the goals of actions are available only where a theory of the evidential basis of mindreading 
 takes the possibility of interaction into account.
In what follows we shall explain how abilities to interact with others provide a route to knowledge of the goals of other agents' actions  which avoids the problem of opaque means.
This is one way in which moving away from a purely observational model of mindreading yields a richer evidential base for ascriptions.

@@

Let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
First, this outcome is a one to which each agent's actions are individually directed.
Second, each agent's actions are related to the outcome in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed in bringing about this outcome.

@@

In defining `distributive goal' we stipulated that it must be possible for all of the agents of actions with a distributive goal to succeed relative to that outcome.
Consider a variation on our first turnip-pulling story.
Instead of each intending that we pull up the turnip,
we each simply intend  to pull up the turnip.
In this case,
some might claim that
this amounts to us each intending that he or she pulls up the turnip.
And this claim might be taken to support the further claim that
there is no outcome to which each of our actions are directed
in virtue of our having these intentions
such that we could all succeed relative to this outcome.
%, given our intentions in this alternative case, we could not all succeed relative to the outcome of pulling up the turnip---%
%that if your actions were to succeed relative to the turnip's extraction,
%then ours would fail.
We take no stand on whether either claim is correct.
For our purposes all that matters is that there are distributive goals.
And the existence of distributive goals follows from the possibility of each of us 
intending that we pull up the turnip
(rather than simply
intending to pull it up).

@@

For two agents' actions to have a distributive goal it is sufficient that their actions constitute a joint action 
(at least this is true on almost any account of joint action).%
\footnote{
We know of two definitions of joint action on which it is not straightforward that all joint actions involve distributive goals.
One is provided 
 is provided by \citet[p.\ 366]{ludwig_collective_2007} and discussed in footnote \vref{fn:ludwig}; the other is due to 
\citet[p.\ 70]{Sebanz:2006yq}.
According to Sebanz and colleagues' `working definition', 
`joint action can be regarded as any form of social interaction whereby two or more individuals coordinate their actions in space and time to bring about a change in the environment'
(\citeyear[p.\ 70]{Sebanz:2006yq}).
On one way of reading this, the latter part is equivalent to `a change in the environment is such that: to bring it about two or more individuals coordinate their actions';
 this does suggest that the actions have a distributive goal.
On other standard accounts it is straightforward that joint action involves distributive goals.
Compare 
	\citet[pp.\ 329-31]{Bratman:1992mi},
%	\citet[p.\ 40]{Butterfill:2011fk},
	\citet[pp.\ 96-7]{Searle:1990em} and
	\citet[pp.\ 168-9]{gilbert:2009shared}.
}
However, the converse does not hold.
As this the turnip-pulling variation  indicates,
two or more agents' actions may have a distributive goal 
even though the agents do not know about each others' intentions or actions.
In fact our actions might have a distributive goal even though none of us is aware of this, or even of the others' existence. 
(We shall refrain from illustrating this point by contriving a truly gigantic turnip and a completely dark, very windy night, but we are sure readers will be able to imagine how this would go.)
One consequence of this is that 
two or more agents' actions may have a distributive goal
even though they are not engaged in joint action
(at least not on any standard account of joint action).%
\footnote{
\label{fn:ludwig}
According to 
	 \citet[p.\ 366]{ludwig_collective_2007}, 
`The concept of a joint action as such is just that of an event of which there are multiple agents.'
Depending on what events are and what it is to be the agent of an event,
it may turn out that,
on Ludwig's definition,
our actions' having a distributive goal 
is sufficient for
us to be engaged in joint action.
However, other accounts of joint action have the consequence that each agent involved in a joint action must believe or know something about the jointness of her action.
See 
	\citet[p.\ 103]{Bratman:1993je}, %common knowledge
	\citet[p.\ 40]{Butterfill:2011fk},
	\citet[p.\ 10]{Kutz:2000si}, %Agents acting on a shared intention have `a conception of themselves as contributors to a collective end.'
	\citet[p. 56]{miller_social_2001} % requires that each agent believes her actions are interdependent with the other agent's.
	and
	\citet[p.\ 361]{Roth:2004ki}. %: `each participant ... can answer the question of what he is doing or will be doing by saying for example ``We are walking together'' or ``We will/intend to walk together.''' 
}

The notion of a distributive goal is a 
narrowly technical one.
In introducing this notion by stipulation we are not aiming to match anyone's intentions about anything.
What matters for now is just that the definition is theoretically coherent and that it is possible for actions to have distributive goals.
Our aim in introducing distributive goals 
 is to capture a minimal requirement on the sort of interactions 
relevant to characterising a route to knowledge of the goals of actions
available only to interacting mindreaders.
% 
% relevant to explaining how 
%interacting mindreaders 
%might be able to exploit routes to knowledge of the goals of actions
%which would be unavailable  were  they mere observers.

\section{Your-goal-is-my-goal: a route to knowledge}
\label{sec:your_goal_is_my_goal}
If a mindreader is able to interact with her targets,
if she is not limited to merely observing them,
how might this provide her with a route to knowledge of the goals of their actions?
The answer hinges on interactions involving distributive goals
(see \vref{sec:joint_action}).

Here is an intuitive idea that doesn't quite work:
if a mindreader is engaged in an interaction with her target that involves a distributive goal, 
it may be easy for the interpreter to know what the goal of her target's actions is because this goal is the goal of her own actions.
So if she knows the goal of her own actions and she knows that she is engaged with her target in an interaction involving a distributive goal,
then she already knows what the goal of her target's actions are.
%Roughly speaking, the interpreter can reason about her target thus: your goal is my goal.

Of course this intuitive idea won't work as it stands.  
For the inference it captures relies on the premise that the mindreader and her target are engaged in actions with a distributive goal.
But for the mindreader to know this premise it seems she must   already know which goal her target's actions are directed to.

Fortunately there is a way around this.  
For there are various cues which signal that one agent is prepared to engage in joint action with another,
and joint actions involve distributive goals.  
\label{twin_pram}
Seeing a stranger with the haggard look of a new parent struggling to get his twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.
(The noise of the street rules out talking, 
and you can barely speak the local language anyway.)   
In this way you signal that you are about to act jointly with the stranger, to lift the pram with him.   
Since the stranger is fully committed to getting his pram onto the bus,
he knows what the sole goal of his own actions will be.
But he also knows that you will engage in joint action with him,
which means that your actions and his actions will have a distributive goal.
This enables him to infer the goal of your imminent actions: 
your goal is his goal, to get the pram onto the bus.

Our suggestion, then, is that the following inference characterises a route to knowledge of others’ goals:
%
\begin{enumerate}
\item You are willing to engage in some joint action\footnote{
We leave open the issue of how joint action is to be characterised subject only to the 
requirement that all joint actions must involve distributive goals.
Attempts to characterise joint action in ways relevant to explaining development include
	\citet{Tollefsen:2005vh},
	\citet{Carpenter:2009wq},
	\citet{pacherie_framing_2011} and
	\citet{Butterfill:2011fk}.
The last of these 
not only shows how a primitive form of joint action can be characterised without ascriptions of higher-order mental states
but also shares our focus on relations between actions and goals rather than between agents and intentions.
}
or other with me
%(for example, because you have made eye contact with me while I was in the middle of attempting to do something).

@@

Not everything needs to rest on indicators, however.
It is sometimes possible to know that others are willing to engage in joint action with you even without relying on such indicators.  
Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are willing to act jointly.  
For example, this is often so for children surrounded by family or familiar adults who are struggling with a coat.  
And in at least some subcultures people using public transport can reasonably take for granted that, within limits, those around them will act jointly with them when the need is clear.
Of course dispositions to engage in joint action may vary between cultural groups and situations.
This may be fatal for the peripatetic cosmopolitan,
but for others what matters is not whether the dispositions are universal but only that they are sufficiently widespread to be predictable.

Turning to the second point,
could knowing the premises of the your-goal-is-my-goal inference ever put one in a position to know the conclusion?  
Of course the inference is not deductive and will only work when certain background conditions are met.  
These background conditions include the other having largely true beliefs concerning which goal your actions are or will be directed to.
After all, the other agent may be willing to act jointly with you while being entirely mistaken about the goals to which your actions will be directed.
Where this happens, the premises of the inference might be true but the conclusion false.
(We return to this point in section \vref{sec:false_belief}.)  
But there are situations in which it is reasonable  to ignore  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
Given the right conditions,
that others are willing to engage in joint action with you 
is sometimes sufficient reason to hold that
you will end up acting jointly with them
 even where the goal to which your actions will be directed is already fixed.

@@

The your-goal-is-my-goal route to knowledge is characterised by an inference.  
However, exploiting this route to knowledge may not require actually making the inference or knowing the premises.  
Depending on what knowing requires,
it may be sufficient to believe the conclusion because one has reliably detected a situation in which the premises of the inference are true;
it may not be necessary to think of this situation as a situation where the premises are true,
nor even to be able to think of it in this way.

In principle,
exploiting your-goal-is-my-goal does not require actually being in a position to interact with the target of goal ascription.
It is sufficient (and would be necessary but for some special cases) to be taken by the target to be in a position to interact. 
Our concern, however,
is with cases that are likely to be important for understanding
development or evolution (and ideally both).
This motivates a focus on goal ascribers of limited sophistication
who lack both deep insight into others' minds and fully-fledged communicative abilities.
Such interpreters are unlikely to be able to contrive or exploit situations in which they only appear to be in a position to interact with their targets.
When we come to applications of your-goal-is-my-goal, the relevant cases will be those 
in which a goal ascriber is manifestly in a position to interact with her target.

Our aim in identifying the your-goal-is-my-goal inference is not to 
defend a detailed hypothesis about  mechanisms and processes supporting mindreading.
Instead our present concerns are limited to a normative question about the evidential basis of mindreading.
The your-goal-is-my-goal inference matters
not because it describes exactly how anyone actually assigns goals (maybe it doesn't)
but because 
it characterises a route to knowledge
that is closed to mere observers
and open to interacting mindreaders.

@@

In our earlier example 
a novice parent is struggling to lift his heavy twin pram onto a bus when a stranger joins in and they lift the pram together (\vpageref*{twin_pram}).
Suppose the stranger starts tipping the pram in a way that the inexperienced parent fails to recognise as a means, indeed the only means, of getting it onto the bus.
Outside the context of joint action this might give the parent sufficient evidence to reject the idea that the goal of the stranger's actions is to get the pram onto this bus.
Or suppose the stranger starts by pulling the pram away from the bus in order to better position it for entry,
and the novice parent does not realise that this is necessary.
Outside the context of joint action, 
his evidence might on balance support the conclusion that the goal of the stranger's actions is to take the pram off the bus.
(Perhaps the stranger is impatient to get onto the bus herself.)
But in the context of joint action,
your-goal-is-my goal gives the parent  additional evidence for supposing that, 
even though the stranger's actions do not seem to him to be a means to getting the pram onto the bus, 
this really is the goal of her actions.
Of course this additional evidence will not necessarily trump   other evidence.
But it does provide a possible way around the problem of opaque means by providing evidence for goal ascriptions that is independent of a goal ascriber's understanding of which ends actions are means to.

@@

As this example indicates,
exploiting your-goal-is-my-goal 
can shift the burden of identifying goals from a mindreader to her target.
In the example Ayesha is the focal mindreader and we are her target;
but her success in identifying the goal of our actions
depends on this,
that our willingness to act jointly with her is based on \emph{our} knowledge of the goals of \emph{her} actions.
In purely observational mindreading, the target's beliefs about the goals of the mindreader's actions are not normally relevant (except, of course, when the mindreader is ascribing such beliefs).
But interacting mindreaders who rely on your-goal-is-my-goal
thereby rely on their targets' having correctly identified the goals of their actions.
Of course this is sometimes a reason not to rely on your-goal-is-my-goal.
But where the target understands relevant means-ends relations,
such as actions involving novel tools,
the your-goal-is-my-goal route to knowledge of other's goals may sometimes be the only option.

@@

\section{Communicative gestures}
\label{sec:communicative_gestures}
When introducing the problem of opaque means 
(in section \vref{sec:opaque_means})
we saw that it could  affect
communicative actions.
To illustrate this suggestion we drew on 
an experiment by \citet{hare_chimpanzees_2004}
in which chimpanzees had to find a reward 
and were helped by being shown either a failed reach or  a helpful point to the target location.
Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which  did help.
Taking this paradigm as a case study, 
we want to suggest that your-goal-is-my-goal might 
enable us to understand how abilities to engage in joint action 
could be part of what enables mindreaders to make the transition from 
a simple understanding of goals 
to 
an early understanding communicative actions.%
\footnote{
Note that although our discussion borrows an experimental paradigm,
our aim is not to argue for empirical hypotheses about chimpanzee social cognition.
Our aim is only to argue for the theoretical significance of interaction for mindreading by showing 
that your-goal-is-my-goal
could \emph{in principle} enable
individuals to make the transition 
from
 a simple understanding of goals
to
 an early understanding of communicative actions.
Of course it would powerfully demonstrate the relevance of our theoretical argument if we could provide evidence to show that this actually happens.
But for now we are concerned with more narrowly conceptual issues.
}

@@

So far we have illustrated how your-goal-is-my-goal enables responding appropriately to communicative gestures with two examples, pointing and holding up a replica.
The pattern of reasoning generalises to a wider range of communicative gestures including single-word utterances.
The basic requirement is this: in a particular context, the goal ascriber must associate a communicative gesture with its referent.
For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking to see who has an object she must associate an utterance of `daddy' with the daddy.%
\footnote{
Such associations appear to be in place by six months of age or earlier \citep{tincoff:1999_beginnings_,tincoff:2011}.
There is debate on when infants are first able to rapidly associate words with novel objects \citep{werker:1998_acquisition,friedrich:2011_word_}.
Interestingly,
a possible developmental decalage separating 
two measures of whether infants have formed a word-object association,
preferential looking and comprehension as measured by reaching in response to a request  \citep{gurteen:2011_rapid},
may suggest that
infants can form a word-object association before they can comprehend communicative actions involving the word.
}
As we saw, 
outside the context of joint action,
merely associating a gesture with its referent falls short of being able to respond appropriately.
But if an interpreter supposes that her target is willing to engage in joint action with her,
then she may infer that the goal of her target's action is her goal
and so be motivated to treat the thing associated with a communicative gesture as relevant to the goal of her own actions.
This will reliably (but not always) enable her to respond appropriately to the communicative gesture even without understanding it as a communicative gesture.
And once she has experienced how that communicative gesture works as a tool for guiding others' actions in the context of joint action,
she may be in a position to realise, further, that the same tool can be used in other contexts.

>>>>

We are not suggesting that no observer could ever identify the goal of any action she fails to recognise as a means to achieving that goal.
Of course opaque means are not in every case an insurmountable obstacle to goal ascription, and
they may only rarely be a problem for human adults with sophisticated social skills.
Our point is neither novel nor surprising:
opaque means sometimes  deprive mindreaders of evidence and so prevent goal ascription.
This is more likely to happen where  goal ascribers lack sophistication in mindreading, communication and culture.

Some of the most plausibly unique aspects of human cognition depend on our abilities to recognise the goals of novel behaviours involving tools, and of communicative gestures.
The problem of opaque means is likely to arise in both cases
if goal ascription is based  entirely on observation 
(so that the possibility of interaction is ignored)
and if the goal ascriber lacks sophisticated social skills.
We have just seen an illustration of how the problem of opaque means arises where tools are used to unfamiliar ends.
Relatedly, it is also likely to arise where actions involve multiple steps that do not form a familiar sequence, can occur in various orders and can be interspersed among other activities;
as in preparing spirit from grain, for example.
%\footnote{
%*cut because 'plausibly unique aspects of human cognition'*
%one example is food preparation in some groups of mountain gorillas
%\citep[p.\ 531]{Byrne:2003wx}.
%*This is slightly tricky because Byrne wants a pure behaviour reading account with no understanding of intention.  
%Stress that this is consistent with supposing that goals are ascribed.
%}

@@

This, then, is the problem of opaque means:
failures to identify to which ends actions are means can impair goal ascription.
The problem is potentially a problem for mindreaders given the standard, purely observational models of mindreading.
It is not our intention to suggest that the problem of opaque means is a problem \emph{for models of mindreading}.
Rather, it is a potential problem \emph{for mindreaders}.

That the problem of opaque means exists
is the first premise in our argument
that models of mindreading based on pure observation are less powerful than models taking into account the possibility of interaction.
They are less powerful in this sense: 
some routes to knowledge of the goals of actions are available only where a theory of the evidential basis of mindreading 
 takes the possibility of interaction into account.
In what follows we shall explain how abilities to interact with others provide a route to knowledge of the goals of other agents' actions  which avoids the problem of opaque means.
This is one way in which moving away from a purely observational model of mindreading yields a richer evidential base for ascriptions.

@@

Let us stipulate that an outcome is a \emph{distributive goal} of two or more agents' actions just if two conditions are met.
\label{df:distributive_goal}
First, this outcome is a one to which each agent's actions are individually directed.
Second, each agent's actions are related to the outcome in such a way that it is possible for all the agents (not just any agent, all of them together) to succeed in bringing about this outcome.

@@

@@

In defining `distributive goal' we stipulated that it must be possible for all of the agents of actions with a distributive goal to succeed relative to that outcome.
Consider a variation on our first turnip-pulling story.
Instead of each intending that we pull up the turnip,
we each simply intend  to pull up the turnip.
In this case,
some might claim that
this amounts to us each intending that he or she pulls up the turnip.
And this claim might be taken to support the further claim that
there is no outcome to which each of our actions are directed
in virtue of our having these intentions
such that we could all succeed relative to this outcome.
%, given our intentions in this alternative case, we could not all succeed relative to the outcome of pulling up the turnip---%
%that if your actions were to succeed relative to the turnip's extraction,
%then ours would fail.
We take no stand on whether either claim is correct.
For our purposes all that matters is that there are distributive goals.
And the existence of distributive goals follows from the possibility of each of us 
intending that we pull up the turnip
(rather than simply
intending to pull it up),
and from the possibility of each of us
intending to contribute to pulling up the turnip or pull it up.

For two agents' actions to have a distributive goal it is sufficient that their actions constitute a joint action 
(at least this is true on almost any account of joint action).%
\footnote{
\label{fn:joint_action_distributive_goal}
We know of two definitions of joint action on which it is not straightforward that all joint actions involve distributive goals.
One is provided 
 is provided by \citet[p.\ 366]{ludwig_collective_2007} and quoted in footnote \vref{fn:ludwig}; the other is due to 
\citet[p.\ 70]{Sebanz:2006yq}.
According to Sebanz and colleagues' `working definition', 
`joint action can be regarded as any form of social interaction whereby two or more individuals coordinate their actions in space and time to bring about a change in the environment'
(\citeyear[p.\ 70]{Sebanz:2006yq}).
On one way of reading this, the latter part is equivalent to `a change in the environment is such that: to bring it about two or more individuals coordinate their actions';
 this does suggest that the actions have a distributive goal.
On other standard accounts it is straightforward that joint action involves distributive goals.
Compare 
	\citet[pp.\ 329-31]{Bratman:1992mi},
	\citet[pp.\ 96-7]{Searle:1990em} and
	\citet[pp.\ 168-9]{gilbert:2009shared}.
}
However, the converse does not hold.
As the first variation of the turnip-pulling story indicates,
two or more agents' actions may have a distributive goal 
even though the agents do not know about each others' intentions or actions.
In fact our actions might have a distributive goal even though none of us is aware of this, or even of the others' existence. 
(We shall refrain from illustrating this point by contriving a truly gigantic turnip and a completely dark, very stormy night; readers can probably guess how this would go.)
One consequence of this is that 
two or more agents' actions may have a distributive goal
even though they are not engaged in joint action
(at least not on any standard account of joint action).%
\footnote{
On many accounts of joint action, each agent involved in a joint action must believe, expect or know something about the jointness of her action.
See 
	\citet[p.\ 103]{Bratman:1993je}, %common knowledge
	\citet[p.\ 40]{Butterfill:2011fk},
	\citet[p.\ 10]{Kutz:2000si}, %Agents acting on a shared intention have `a conception of themselves as contributors to a collective end.'
	\citet[p. 56]{miller_social_2001} % requires that each agent believes her actions are interdependent with the other agent's.
	and
	\citet[p.\ 361]{Roth:2004ki}. %: `each participant ... can answer the question of what he is doing or will be doing by saying for example ``We are walking together'' or ``We will/intend to walk together.''' 
Exceptions include  \citet{pacherie_framing_2011} and \citet{ludwig_collective_2007}.
\label{fn:ludwig}
According to Ludwig,
`The concept of a joint action as such is just that of an event of which there are multiple agents' (\citeyear[p.\ 366]{ludwig_collective_2007}).
Depending on what events are and what it is to be the agent of an event,
it may turn out that,
on Ludwig's definition,
our actions' having a distributive goal 
is sufficient for
us to be engaged in joint action.
}

The notion of a distributive goal is a 
narrowly technical one.
In introducing this notion by stipulation we are not aiming to match anyone's intuitions about anything.
What matters for now is just that the definition is theoretically coherent and that it is possible for actions to have distributive goals.
Our aim in introducing distributive goals 
 is to capture a minimal requirement on the sort of interactions 
relevant to characterising a route to knowledge of the goals of actions
available only to interacting mindreaders.
% 
% relevant to explaining how 
%interacting mindreaders 
%might be able to exploit routes to knowledge of the goals of actions
%which would be unavailable  were  they mere observers.

\section{Your-goal-is-my-goal}
\label{sec:your_goal_is_my_goal}
If a mindreader is able to interact with her targets,
if she is not limited to merely observing them,
how might this provide her with a route to knowledge of the goals of their actions?
The answer hinges on interactions involving distributive goals
(as defined \vpageref{df:distributive_goal}).

Here is an intuitive idea that doesn't quite work:
if a mindreader is engaged in an interaction with her target that involves a distributive goal, 
it may be easy for the mindreader to know what the goal of her target's actions is because this goal is the goal of her own actions.
So if she knows the goal of her own actions and she knows that she is engaged with her target in an interaction involving a distributive goal,
then she already knows what the goal of her target's actions are.
%Roughly speaking, the mindreader can reason about her target thus: your goal is my goal.

Of course this intuitive idea is no use it stands.  
For the inference it captures relies on the premise that the mindreader and her target are engaged in actions with a distributive goal.
But for the mindreader to know this premise it seems she must   already know which goal her target's actions are directed to.

Fortunately there is a way around this.  
For there are various cues which signal that one agent is prepared to engage in joint action with another,
and joint actions involve distributive goals.  
\label{twin_pram}
Seeing a stranger with the haggard look of a new parent struggling to get his twin pram on to the bus, you grab the front wheels and make eye contact, raising your eyebrows and smiling.
(The noise of the street rules out talking.)   
In this way you signal that you are about to act jointly with the stranger, to lift the pram with him.   
Since the stranger is fully committed to getting his pram onto the bus,
he knows what the sole goal of his own actions will be.
But he also knows that you will engage in joint action with him,
which means that your actions and his actions will have a distributive goal.
This may enable him to infer the goal of your imminent actions: 
your goal is his goal, to get the pram onto the bus.

Our suggestion, then, is that the following inference characterises a route to knowledge of others’ goals:
%
\begin{enumerate}
\item You are willing to engage in some joint action\footnote{
We leave open the issue of how joint action is to be characterised subject only to the 
requirement that all joint actions must involve distributive goals (see further footnote \vref{fn:joint_action_distributive_goal}).
Attempts to characterise joint action in ways relevant to explaining development include
	\citet{Tollefsen:2005vh},
	\citet{Carpenter:2009wq},
	\citet{pacherie_framing_2011} and
	\citet{Butterfill:2011fk}.
The last of these 
not only shows how a primitive form of joint action can be characterised without ascriptions of higher-order mental states
but also shares our focus on relations between actions and goals rather than between agents and intentions.
}
or other with me
%(for example, because you have made eye contact with me while I was in the middle of attempting to do something).

@@

Not everything needs to rest on indicators, however.
It is sometimes possible to know that others are willing to engage in joint action with you even without relying on such indicators.  
Thanks to widespread dispositions to act jointly, in some situations it is reasonable to take for granted that others are willing to act jointly.  
For example, this is often so for children struggling with a coat while surrounded by family or familiar adults.  
And in at least some subcultures people using public transport can reasonably take for granted that, within limits, those around them will act jointly with them when the need is clear.
Of course dispositions to engage in joint action may vary between cultural groups and situations.
This may be fatal for the peripatetic cosmopolitan,
but for others what matters is not whether the dispositions are universal but only that they are sufficiently widespread to be predictable.

Turning to the second point,
could knowing the premises of the your-goal-is-my-goal inference ever put one in a position to know the conclusion?  
Of course the inference is not deductive and will only work when certain background conditions are met.  
These background conditions include the other having largely true beliefs concerning which goal your actions are or will be directed to.
After all, the other agent may be willing to act jointly with you while being entirely mistaken about the goals to which your actions will be directed.
Where this happens, the premises of the inference might be true but the conclusion false.
(We return to this point in section \vref{sec:false_belief}.)  
But in some situations there is no requirement  to consider  this possibility, as for instance when stereotypes, conventions or simplicity should and do make the goal of your actions obvious to the other agent.
Given the right conditions,
that others are willing to engage in joint action with you 
is sometimes sufficient reason to hold that
you will end up acting jointly with them
 even where the goal to which your actions will be directed is already fixed.

@@

In principle,
exploiting your-goal-is-my-goal does not require actually being in a position to interact with the target of goal ascription.
It is sufficient (and would be necessary but for some special cases) to be taken by the target to be in a position to interact. 
Our concern, however,
is with cases that are likely to be important for understanding
development or evolution.
This motivates a focus on goal ascribers of limited sophistication
who lack both deep insight into others' minds and fully-fledged communicative abilities.
Such individuals are unlikely to be able to contrive or exploit situations in which they only appear to be in a position to interact with their targets.
When we come to applications of your-goal-is-my-goal, the relevant cases will be those 
in which a goal ascriber is manifestly in a position to interact with her target.

The your-goal-is-my-goal route to knowledge is characterised by an inference.  
However, exploiting this route to knowledge may not require actually making the inference or knowing the premises.  
Depending on what knowing requires,
it may be sufficient to believe the conclusion because one has reliably detected a situation in which the premises of the inference are true;
it may not be necessary to think of this situation as a situation where the premises are true,
nor even to be able to think of it in this way.
Our aim in identifying the your-goal-is-my-goal inference is not to 
defend a detailed hypothesis about  mechanisms of mindreading.
Instead our present concerns are limited to a normative question about the evidential basis of mindreading.
The your-goal-is-my-goal inference matters
not because it describes exactly how anyone actually assigns goals (maybe it doesn't)
but because 
it characterises a route to knowledge
that is closed to mere observers
and open to interacting mindreaders.

@@

In our earlier example 
a novice parent is struggling to lift his heavy twin pram onto a bus when a stranger joins in and they lift the pram together (\vpageref*{twin_pram}).
Suppose the stranger starts tipping the pram in a way that the inexperienced parent fails to recognise as a means, indeed the only means, of getting it onto the bus.
Outside the context of joint action, 
failure to realise that the action is a means to getting the pram onto the bus
might have the consequence that
the parent's evidence on balance supports the conclusion that the goal of the stranger's actions is to take the pram off the bus.
(Perhaps the stranger is impatient to get onto the bus herself.)
%Or suppose the stranger starts by pulling the pram away from the bus in order to better position it for entry,
%and the novice parent does not realise that this is necessary.
%Outside the context of joint action, 
%his evidence might on balance support the conclusion that the goal of the stranger's actions is to take the pram off the bus.
%(Perhaps the stranger is impatient to get onto the bus herself.)
But in the context of joint action,
your-goal-is-my goal gives the parent  additional evidence for supposing that, 
even though the stranger's actions do not seem to him to be a means to getting the pram onto the bus, 
this really is the goal of her actions.
Of course this additional evidence will not always outweigh   other evidence.
But it will do sometimes, and this is all we need.
We have identified evidence for goal ascription that is available independently of a goal ascriber's knowing which ends actions are means to, and we have illustrated how this will sometimes enable mindreaders to escape the problem of opaque means

@@

As this example indicates,
exploiting your-goal-is-my-goal 
can shift the burden of identifying goals from a mindreader to her target.
In the example Ayesha is the focal mindreader and we are her target;
but her success in identifying the goal of our actions
depends on this,
that our willingness to act jointly with her is based on \emph{our} knowledge of the goals of \emph{her} actions.
In purely observational mindreading, the target's beliefs about the goals of the mindreader's actions are not normally relevant (except, of course, when the mindreader is ascribing such beliefs).
But interacting mindreaders who rely on your-goal-is-my-goal
thereby rely on their targets' having correctly identified the goals of their actions.
Of course this is sometimes a reason not to rely on your-goal-is-my-goal.
But where the target understands relevant means-ends relations,
such as actions involving novel tools,
the your-goal-is-my-goal route to knowledge of others' goals may sometimes be the only option.

Here we have a first illustration of how
interacting mindreaders can exploit reciprocity 
without relying on ascriptions of higher-order mental states.
The mindreader's knowledge of the goal of her target's actions depends on the target's knowledge of the goal of the mindreader's actions.
What makes this dependence possible is not each knowing that the other knows something about the goal of her actions;
 it is  the actual or apparent  possibility of interaction.

@@

\section{Communicative gestures}
\label{sec:communicative_gestures}
When introducing the problem of opaque means 
(in section \vref{sec:opaque_means})
we saw that it could  affect
communicative actions.
To illustrate this suggestion we drew on 
an experiment by \citet{hare_chimpanzees_2004}
in which chimpanzees had to find a reward 
and were helped by being shown either a failed reach or  a helpful point to the target location.
Strikingly, for chimpanzees the helpful point is no help at all---even though it superficially resembles the failed reach, which  did help.
Taking this paradigm as a case study, 
we want to suggest that your-goal-is-my-goal might 
enable us to understand how abilities to engage in joint action 
could be part of what enables mindreaders to make the transition from 
a simple understanding of goals 
to 
an early understanding communicative actions.%
\footnote{
Note that although our discussion borrows an experimental paradigm,
our aim is not to argue for empirical hypotheses about chimpanzee social cognition.
Our aim is only to argue for the theoretical significance of interaction for mindreading by showing 
that your-goal-is-my-goal
could {in principle} enable
individuals to make the transition 
from
 a simple understanding of goals
to
 an early understanding of communicative actions.
Of course it would powerfully demonstrate the relevance of our theoretical argument if we could provide evidence to show that this actually happens.
But for now we are concerned with more narrowly conceptual issues.
}

@@

So far we have illustrated how your-goal-is-my-goal enables responding appropriately to communicative gestures with two examples, pointing and holding up a replica.
The pattern of reasoning generalises to a wider range of communicative gestures including single-word utterances.
The basic requirement is this: in a particular context, the goal ascriber must associate a communicative gesture with its referent.
For instance, she must associate the pointing gesture with the object indicated; or, if (say) she is looking to see who has an object she must associate an utterance of `daddy' with the daddy.%
\footnote{
Such associations appear to be in place by six months of age or earlier \citep{tincoff:1999_beginnings_,tincoff:2011}.
There is debate on when infants are first able to rapidly associate words with novel objects \citep{werker:1998_acquisition,friedrich:2011_word_}.
Interestingly,
a possible developmental decalage separating 
two measures of whether infants have formed a word-object association,
preferential looking and comprehension as measured by reaching in response to a request  \citep{gurteen:2011_rapid},
may suggest that
infants can form a word-object association before they can comprehend communicative actions involving the word.
}
As we saw, 
outside the context of joint action,
merely associating a gesture with its referent falls short of being able to respond appropriately.
But if a mindreader supposes that her target is willing to engage in joint action with her,
then she may infer that the goal of her target's action is her goal
and so be motivated to treat the thing associated with a communicative gesture as relevant to the goal of her own actions.
This will reliably (but not always) enable her to respond appropriately to the communicative gesture even without understanding it as a communicative gesture.
And once she has experienced how that communicative gesture works as a tool for guiding others' actions in the context of joint action,
she may be in a position to realise, further, that the same tool can be used in other contexts.

This, in barest outline, is 
how
possessing abilities to engage in joint action
means that 
an individual with an ability to ascribe simple goals only and no understanding of communicative intent
might 
nevertheless reliably respond appropriately to some communicative gestures,
and so come be in a position to understand how such gestures can be used to guide others' actions.%
\footnote{
Contrast Csibra's claim that,
early in human development,
goal ascription (`teleological understanding' in his terms)
and 
identifying the referents of communicative gestures (`referential understanding')
`rely on different kinds of action understanding' %\citep[p.\ 456]{Csibra:2003kp}
and
are initially two distinct `action interpretation systems'   \citep[p.\ 456]{Csibra:2003kp}.
%(although of course they come together later in development)
We have not shown that this claim wrong.
But we have shown that there is another possibility:
the referential stance may 
emerge from
the teleological stance
 together with abilities to engage in simple forms of joint action.
}
Of course we have only argued that this transition is theoretically possible;
we have not attempted to defend any hypothesis about 
anything's actual evolution or development.
%Our point is just to illustrate one of several ways in which interaction may matter for mindreading.
But widening the evidential basis for goal ascription
to include evidence available only to mindreaders actually or apparently capable of interacting with their targets
does reveal the coherence of  
novel hypotheses about a role for interaction in the emergence  of communicative gestures.
